---
title: "Paths of Change: Community Trajectory Aanalysis of Everglades fish assemblages over 20 years"
authors: "James Sturges"
date-modified: 30 July 2025
format: html
editor: visual
format: html
editor: visual
---

## **Section 1: Data Wrangling**

##### - Required Packages

```{r}
#| warning: false
#| message: false
library(raster) 
library(gdm)            
library(reshape2)         
library(lme4)             
library(MASS)             
library(AER)
library(vegan)              
library(ecotraj)
library(purrr)
library(RColorBrewer)
library(vegclust)
library(ecotraj)
library(smacof)
library(tidyverse)
library(lubridate)
library(remotes)
library(devtools)
```

##### - Raw Data Availability

```{r}
#MAP master file 
raw_data = read.csv("data/MAP_master.csv")

```

##### - Site Metadata

```{r}
#| warning: false
#site metadata (15 long-term sites with 3, 100m bouts of electrofishing)
site_metadata <- read.csv("data/map_waypoints.csv")
```

##### - Species Metadata

```{r}
species_metadata = read.csv("data/spp_codes.csv")
```

##### - Joining Species Metadata

```{r}
# Join the dataframes
raw_data <- raw_data %>%
  mutate(code = as.character(code))
# 
species_metadata <- species_metadata %>%
  mutate(code = as.character(code))

MAP <- raw_data %>%
  left_join(species_metadata, by = "code")
```

##### - Joining Site Metadata

```{r}
MAP = MAP %>% 
  mutate(drainage.site = paste0(drainage,site))

MAP = MAP %>% 
  mutate(drainage.site.bout = paste0(drainage,site,"B",bout))

MAP = MAP %>% 
  mutate(drainage.site.year = paste0(drainage,site,"-", year))

site_metadata <- site_metadata %>%
      rename("drainage.site.bout" = "Site.Name")

```

##### - Standardize Sampling Effort

```{r}
core = c("RB8", "RB9", "RB10", "RB11", "RB13")

MAP_core_sites <- MAP %>%
  filter(drainage.site %in% core)


# Count number of unique months sampled at each site-year combination
site_year_month_summary <- MAP_core_sites %>%
  distinct(site, year, month) %>%
  group_by(site, year) %>%
  summarise(n_months = n_distinct(month), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = n_months, values_fill = 0)


site_month_check <- MAP_core_sites %>%
  distinct(drainage.site.year, month) %>%
  mutate(sampled = 1) %>%
  pivot_wider(names_from = month, values_from = sampled, values_fill = 0)


# Generate the presence table
presence_matrix <- MAP_core_sites %>%
  distinct(site, year, month) %>%
  mutate(present = 1) %>%
  complete(site, year, month, fill = list(present = 0))  # Ensure all combos exist

# Count number of sites that have a given year × month
site_counts <- presence_matrix %>%
  group_by(year, month) %>%
  summarise(n_sites_sampled = sum(present), .groups = "drop")

# Total number of unique sites
n_sites_total <- MAP_core_sites %>% pull(site) %>% unique() %>% length()

# Filter to only year-months sampled by all sites
complete_year_months <- site_counts %>%
  filter(n_sites_sampled == n_sites_total)

# Use the complete_year_months
decision_table <- complete_year_months %>%
  mutate(sampled = 1) %>%
  pivot_wider(
    names_from = month,
    values_from = sampled,
    values_fill = 0
  ) %>%
  arrange(desc(year)) 

years_to_keep <- c("7", "9", "10", "13", "14", "15", "17", "18", "20")  
months_to_keep <- c("DECEMBER", "FEBRUARY", "MARCH", "APRIL", "MAY")

decision_table <- decision_table |>
  select(-JANUARY, -NOVEMBER, -SEPTEMBER, -JUNE, -JULY, -AUGUST)

decision_table <- decision_table |>
  mutate(year = as.character(year))

decision_table <- decision_table |>
  filter(year %in% years_to_keep)



MAP_core_sites <- MAP_core_sites |>
  mutate(year = as.numeric(year))

MAP_core_sites_May = MAP_core_sites |> 
  filter(month == "MAY")
MAP_core_sites_April = MAP_core_sites |> 
  filter(month == "APRIL")
MAP_core_sites_December = MAP_core_sites |> 
  filter(month == "DECEMBER")


MAP_core_sites = MAP_core_sites|>
  filter(year %in% years_to_keep)

MAP_core_sites = MAP_core_sites|>
  filter(month %in% months_to_keep)



```

##### - Community Assemblages

```{r}
all_spp = c("American_eel", "Bluegill", "Redear", "Sunfishes", "Spotted_sunfish", "Dollar_sunfish", "Sailfin_molly",
         "Striped_mojarra", "Striped_mullet", "Hogchoker", "Tidewater_mojarra", "Bluespotted_sunfish", "Warmouth","Florida_gar", "Largemouth_bass", "Snook", "Bowfin", "Redfish", "Gray_snapper", "Mangrove_snapper", "Tarpon","Peacock_eel","Blue_tilapia", "Mayan_cichlid", "Jeweled_cichlid")


predator = c("American_eel","Florida_gar", "Largemouth_bass", "Snook", "Bowfin", "Redfish", "Gray_snapper", "Mangrove_snapper", "Tarpon")
  
prey = c("Bluegill", "Redear", "Sunfishes", "Spotted_sunfish", "Dollar_sunfish", "Sailfin_molly",
         "Striped_mojarra", "Striped_mullet", "Hogchoker", "Tidewater_mojarra", "Bluespotted_sunfish", "Warmouth")

invasive = c("Peacock_eel","Blue_tilapia", "Mayan_cichlid", "Jeweled_cichlid")

MAP_core_sites = MAP_core_sites|>
  filter(common_name %in% all_spp)

MAP_core_sites_May = MAP_core_sites_May|>
  filter(common_name %in% all_spp)
MAP_core_sites_December = MAP_core_sites_December|>
  filter(common_name %in% all_spp)
MAP_core_sites_April = MAP_core_sites_April|>
  filter(common_name %in% all_spp)






df_all_spp <- MAP_core_sites %>%
  filter(common_name %in% all_spp)

df_predator <- MAP_core_sites %>%
  filter(common_name %in% predator)

df_prey <- MAP_core_sites %>%
  filter(common_name %in% prey)

df_invasive <- MAP_core_sites %>%
  filter(common_name %in% invasive)
```

##### - Site Color Vectors

```{r}
site_colors <- c(
  "RB8" = "#1b9e77",
  "RB9" = "#d95f02",
  "RB10" = "#7570b3",
  "RB11" = "#e7298a",
  "RB13" = "#1f78b4"
)
```

## Section 2: Matrix Generation

##### - Estimate Biomass Function

```{r}
fill_missing_weights <- function(df) {
  df <- df %>%
    mutate(weight_kg = as.numeric(weight_kg))  # Ensure numeric
  
  # 1. Species × Site × Year mean
  species_site_year <- df %>%
    group_by(species_names, drainage.site, year) %>%
    summarise(mean_w1 = mean(weight_kg, na.rm = TRUE), .groups = "drop")
  
  # 2. Species × Year mean
  species_year <- df %>%
    group_by(species_names, year) %>%
    summarise(mean_w2 = mean(weight_kg, na.rm = TRUE), .groups = "drop")
  
  # 3. Species global mean
  species_overall <- df %>%
    group_by(species_names) %>%
    summarise(mean_w3 = mean(weight_kg, na.rm = TRUE), .groups = "drop")
  
  # Join fallback means to main dataframe
  df_filled <- df %>%
    left_join(species_site_year, by = c("species_names", "drainage.site", "year")) %>%
    left_join(species_year, by = c("species_names", "year")) %>%
    left_join(species_overall, by = "species_names") %>%
    mutate(
      weight_filled = case_when(
        !is.na(weight_kg) ~ weight_kg,
        !is.na(mean_w1) ~ mean_w1,
        !is.na(mean_w2) ~ mean_w2,
        !is.na(mean_w3) ~ mean_w3,
        TRUE ~ NA_real_
      )
    ) %>%
    select(-mean_w1, -mean_w2, -mean_w3)
  
  return(df_filled)
}
```

##### - Fill Missing Biomass

```{r}
MAP_core_sites_filled <- fill_missing_weights(MAP_core_sites)

MAP_core_sites_filled_May <- fill_missing_weights(MAP_core_sites_May)

MAP_core_sites_filled_April <- fill_missing_weights(MAP_core_sites_April)

MAP_core_sites_filled_December <- fill_missing_weights(MAP_core_sites_December)
```

##### - Count Matrix

```{r}
count_matrix_all <- df_all_spp %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = species_names, values_from = count, values_fill = 0)

count_matrix_all <- count_matrix_all %>%
  column_to_rownames("drainage.site.year")

count_matrix_predator <- df_predator %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = species_names, values_from = count, values_fill = 0)

count_matrix_predator <- count_matrix_predator %>%
  column_to_rownames("drainage.site.year")

count_matrix_prey <- df_prey %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = species_names, values_from = count, values_fill = 0)

count_matrix_prey <- count_matrix_prey %>%
  column_to_rownames("drainage.site.year")

count_matrix_invasive <- df_invasive %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = species_names, values_from = count, values_fill = 0)

count_matrix_invasive <- count_matrix_invasive %>%
  column_to_rownames("drainage.site.year")
```

##### - Biomass Matrix

```{r}
biomass_matrix_all <- MAP_core_sites_filled %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(biomass_kg = sum(weight_filled, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = species_names,
    values_from = biomass_kg,
    values_fill = 0
  )

biomass_matrix_all <- biomass_matrix_all %>%
  column_to_rownames("drainage.site.year")


biomass_matrix_predator <- MAP_core_sites_filled %>%
     filter(common_name %in% predator) |> 
  group_by(drainage.site.year, species_names) %>%
  summarise(biomass_kg = sum(weight_filled, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = species_names,
    values_from = biomass_kg,
    values_fill = 0
  )

biomass_matrix_predator <- biomass_matrix_predator %>%
  column_to_rownames("drainage.site.year")


biomass_matrix_prey <- MAP_core_sites_filled %>%
  filter(common_name %in% prey) |> 
  group_by(drainage.site.year, species_names) %>%
  summarise(biomass_kg = sum(weight_filled, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = species_names,
    values_from = biomass_kg,
    values_fill = 0
  )

biomass_matrix_prey <- biomass_matrix_prey %>%
  column_to_rownames("drainage.site.year")


biomass_matrix_invasive <- MAP_core_sites_filled %>%
  filter(common_name %in% invasive) |> 
  group_by(drainage.site.year, species_names) %>%
  summarise(biomass_kg = sum(weight_filled, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = species_names,
    values_from = biomass_kg,
    values_fill = 0
  )

biomass_matrix_invasive <- biomass_matrix_invasive %>%
  column_to_rownames("drainage.site.year")
```

##### - Biomass Matrix Monthly

```{r}
biomass_matrix_all_May <- MAP_core_sites_filled_May %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(biomass_kg = sum(weight_filled, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = species_names,
    values_from = biomass_kg,
    values_fill = 0
  )

biomass_matrix_all_May <- biomass_matrix_all_May %>%
  column_to_rownames("drainage.site.year")


biomass_matrix_all_April <- MAP_core_sites_filled_April %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(biomass_kg = sum(weight_filled, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = species_names,
    values_from = biomass_kg,
    values_fill = 0
  )

biomass_matrix_all_April <- biomass_matrix_all_April %>%
  column_to_rownames("drainage.site.year")


biomass_matrix_all_December <- MAP_core_sites_filled_December %>%
  group_by(drainage.site.year, species_names) %>%
  summarise(biomass_kg = sum(weight_filled, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = species_names,
    values_from = biomass_kg,
    values_fill = 0
  )

biomass_matrix_all_December <- biomass_matrix_all_December %>%
  column_to_rownames("drainage.site.year")


```

## Section 3: CTA Ordinations

##### - Count x All Species

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_all <- vegdist(count_matrix_all, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(count_matrix_all),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

x <- defineTrajectories(
  d = dist_bray_all,
  sites = meta_from_rownames$site,
  times = as.numeric(meta_from_rownames$year)
)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/all_spp_count_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(x,
  d = dist_bray_all,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_all <- trajectoryLengths(x)
lengths_all_df <- lengths_all %>%
  rownames_to_column(var = "site")

# Angles 
angles_all <- trajectoryAngles(x)
angles_all_df <- angles_all %>%
  rownames_to_column(var = "site")

# Convergence/Divergence
convergence_all = trajectoryConvergence(x)



# Export each metric
write_csv(lengths_all_df, "tables/lengths_all_df.csv")
write_csv(angles_all_df, "tables/angles_all_df.csv")
write.csv(convergence_all, "tables/convergence_all.csv")
```

##### - Count x Predator

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_predator <- vegdist(count_matrix_predator, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(count_matrix_predator),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/predator_count_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_predator,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_predator <- trajectoryLengths(d = dist_bray_predator, sites = sites, surveys = years)

# Angles 
angles_predator <- trajectoryAngles(d = dist_bray_predator, sites = sites, surveys = years)

# Convergence/Divergence
convergence_predator = trajectoryConvergence(d = dist_bray_predator, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_predator, path = "tables/lengths_predator.csv")
write_csv(angles_predator, path = "tables/angles_predator.csv")
write.csv(convergence_predator, "tables/convergence_predator.csv")
```

##### - Count x Prey

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_prey <- vegdist(count_matrix_prey, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(count_matrix_prey),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/prey_count_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_prey,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_prey <- trajectoryLengths(d = dist_bray_prey, sites = sites, surveys = years)

# Angles 
angles_prey <- trajectoryAngles(d = dist_bray_prey, sites = sites, surveys = years)

# Convergence/Divergence
convergence_prey = trajectoryConvergence(d = dist_bray_prey, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_prey, path = "tables/lengths_prey.csv")
write_csv(angles_prey, path = "tables/angles_prey.csv")
write.csv(convergence_prey, "tables/convergence_prey.csv")
```

##### - Count x Invasive

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_invasive <- vegdist(count_matrix_invasive, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(count_matrix_invasive),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/invasive_count_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_invasive,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_invasive <- trajectoryLengths(d = dist_bray_invasive, sites = sites, surveys = years)

# Angles 
angles_invasive <- trajectoryAngles(d = dist_bray_invasive, sites = sites, surveys = years)

# Convergence/Divergence
convergence_invasive = trajectoryConvergence(d = dist_bray_invasive, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_invasive, path = "tables/lengths_invasive.csv")
write_csv(angles_invasive, path = "tables/angles_invasive.csv")
write.csv(convergence_invasive, "tables/convergence_invasive.csv")
```

##### -Biomass x All Species

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_all_bio <- vegdist(biomass_matrix_all, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(biomass_matrix_all),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/all_spp_biomass_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_all_bio,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_all_bio <- trajectoryLengths(d = dist_bray_all_bio, sites = sites, surveys = years)

# Angles 
angles_all_bio <- trajectoryAngles(d = dist_bray_all_bio, sites = sites, surveys = years)

# Convergence/Divergence
convergence_all_bio = trajectoryConvergence(d = dist_bray_all_bio, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_all_bio, path = "tables/lengths_all_bio.csv")
write_csv(angles_all_bio, path = "tables/angles_all_bio.csv")
write.csv(convergence_all_bio, "tables/convergence_all_bio.csv")
```

##### -Biomass x All Spp - May

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_all_bio_May <- vegdist(biomass_matrix_all_May, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(biomass_matrix_all_May),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/all_spp_biomass_trajectory_plot_May.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_all_bio_May,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_all_bio_May <- trajectoryLengths(d = dist_bray_all_bio_May, sites = sites, surveys = years)

# Angles 
angles_all_bio_May <- trajectoryAngles(d = dist_bray_all_bio_May, sites = sites, surveys = years)

# Convergence/Divergence
convergence_all_bio_May = trajectoryConvergence(d = dist_bray_all_bio_May, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_all_bio_May, path = "tables/lengths_all_bio_May.csv")
write_csv(angles_all_bio_May, path = "tables/angles_all_bio_May.csv")
write.csv(convergence_all_bio_May, "tables/convergence_all_bio_May.csv")
```

##### -Biomass x All Spp - April

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_all_bio_April <- vegdist(biomass_matrix_all_April, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(biomass_matrix_all_April),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/all_spp_biomass_trajectory_plot_April.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_all_bio_April,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_all_bio_April <- trajectoryLengths(d = dist_bray_all_bio_April, sites = sites, surveys = years)

# Angles 
angles_all_bio_April <- trajectoryAngles(d = dist_bray_all_bio_April, sites = sites, surveys = years)

# Convergence/Divergence
convergence_all_bio_April = trajectoryConvergence(d = dist_bray_all_bio_April, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_all_bio_April, path = "tables/lengths_all_bio_April.csv")
write_csv(angles_all_bio_April, path = "tables/angles_all_bio_April.csv")
write.csv(convergence_all_bio_April, "tables/convergence_all_bio_April.csv")
```

##### -Biomass x All Spp - December

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_all_bio_December <- vegdist(biomass_matrix_all_December , method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(biomass_matrix_all_December ),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/all_spp_biomass_trajectory_plot_December .png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_all_bio_December ,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_all_bio_December  <- trajectoryLengths(d = dist_bray_all_bio_December, sites = sites, surveys = years)

# Angles 
angles_all_bio_December <- trajectoryAngles(d = dist_bray_all_bio_December, sites = sites, surveys = years)

# Convergence/Divergence
convergence_all_bio_December = trajectoryConvergence(d = dist_bray_all_bio_December , sites = sites, surveys = years)



# Export each metric
write_csv(lengths_all_bio_December, path = "tables/lengths_all_bio_December.csv")
write_csv(angles_all_bio_December, path = "tables/angles_all_bio_May.csv")
write.csv(convergence_all_bio_December, "tables/convergence_all_bio_December.csv")
```

##### -Biomass x Predator

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_predator_bio <- vegdist(biomass_matrix_predator, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(biomass_matrix_predator),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/predator_biomass_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_predator_bio,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_predator_bio <- trajectoryLengths(d = dist_bray_predator_bio, sites = sites, surveys = years)

# Angles 
angles_predator_bio <- trajectoryAngles(d = dist_bray_predator_bio, sites = sites, surveys = years)

# Convergence/Divergence
convergence_predator_bio = trajectoryConvergence(d = dist_bray_predator_bio, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_predator_bio, path = "tables/lengths_predator_bio.csv")
write_csv(angles_predator_bio, path = "tables/angles_predator_bio.csv")
write.csv(convergence_predator_bio, "tables/convergence_predator_bio.csv")
```

##### -Biomass x Prey

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_prey_bio <- vegdist(biomass_matrix_prey, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(biomass_matrix_prey),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/prey_biomass_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_prey_bio,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_prey_bio <- trajectoryLengths(d = dist_bray_prey_bio, sites = sites, surveys = years)

# Angles 
angles_prey_bio <- trajectoryAngles(d = dist_bray_prey_bio, sites = sites, surveys = years)

# Convergence/Divergence
convergence_prey_bio = trajectoryConvergence(d = dist_bray_prey_bio, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_prey_bio, path = "tables/lengths_prey_bio.csv")
write_csv(angles_prey_bio, path = "tables/angles_prey_bio.csv")
write.csv(convergence_prey_bio, "tables/convergence_prey_bio.csv")
```

##### -Biomass x Invasive

```{r}
# ---- 1. Create Bray-Curtis Dissimilarity Matrix ----
dist_bray_invasive_bio <- vegdist(biomass_matrix_invasive, method = "bray")

# ---- 2. Extract metadata from rownames ----
meta_from_rownames <- data.frame(
  drainage_site_year = rownames(biomass_matrix_invasive),
  stringsAsFactors = FALSE
) %>%
  separate(drainage_site_year, into = c("site", "year"), sep = "-", remove = FALSE)

# Define grouping (site) and time (year) vectors
sites <- meta_from_rownames$site
years <- as.numeric(meta_from_rownames$year)

# Optional: Define colors for each site
site_colors <- c("RB8" = "#1b9e77",
                 "RB9" = "#d95f02",
                 "RB10" = "#7570b3",
                 "RB11" = "#e7298a",
                 "RB13" = "#1f78b4")

# ---- 3. Generate CTA Plot ----
# Create plot and immediately save to file
png(filename = "figures/CTA/invasive_biomass_trajectory_plot.png", width = 8, height = 8, res = 300, units = "in")

trajectoryPCoA(
  d = dist_bray_invasive_bio,
  sites = sites,
  surveys = years,
  survey.labels = T,
  traj.colors = site_colors,
  lwd = 2
)

legend("topleft", bty = "n",
       legend = names(site_colors),
       col = site_colors,
       lwd = 2)

dev.off()


# ---- 1. Compute CTA Metrics ----
# Lengths
lengths_invasive_bio <- trajectoryLengths(d = dist_bray_invasive_bio, sites = sites, surveys = years)

# Angles 
angles_invasive_bio <- trajectoryAngles(d = dist_bray_invasive_bio, sites = sites, surveys = years)

# Convergence/Divergence
convergence_invasive_bio = trajectoryConvergence(d = dist_bray_invasive_bio, sites = sites, surveys = years)



# Export each metric
write_csv(lengths_invasive_bio, path = "tables/lengths_invasive_bio.csv")
write_csv(angles_invasive_bio, path = "tables/angles_invasive_bio.csv")
write.csv(convergence_invasive_bio, "tables/convergence_invasive_bio.csv")
```
